Containers and R---the Rockerverse and beyond

### 2019-11-06 manuscript.Rmd
one sentence per line
restructure contribution by @karthik, minor changes, add reference

## Common or public work environments [div]

- Binder

The [Binder project](https://mybinder.readthedocs.io/en/latest/), maintained by the team behind Jupyter, makes it possible for users to create and share their computing environments with others. A deployed version of this service called binder hub allows anyone with access to a web browser and an internet connection to launch a temporary instance of these custom environments and execute any notebooks contained within. From a reproducibility standpoint, Binder makes it exceedingly easy to compile a paper, visualize data, and run small examples from papers or tutorials without the need for any local installation.

To set up Binder for a project, a user typically starts at an instance of a Binderhub (examples include mybinder.org and binder.pangeo.io) and passes the location of a hosted Git repository. Using internal tools such as [repo2docker](https://repo2docker.readthedocs.io/en/latest/config_files.html), the binderhub builds a Dockerfile by parsing the contents of all code contained within. While this approach works well for most run of the mill Python projects, it is not so seamless for R projects. For any R projects that use the Tidyverse suite [citation in review. Will add by the time this gets submitted], the time and resources required to build all dependencies from source can often time out before completion, making it frustrating for the average R user.

[`holepunch`](https://github.com/karthik/holepunch) is a R package that was designed to remove some of these limitations and make Binder more accessible to novice R users. `holepunch` achieves this by leveraging Rocker images that contains the Tidyverse along special Jupyter dependencies, and only installs additional packages from CRAN and Bioconductor that are not already part of these images. Starting with the Binder/Tidyverse base images eliminates a large part of the build time and in most cases results in a binder instance launching within a minute. `holepunch` as a side effect also creates a DESCRIPTION file which then turns any project into a research compendium [doi:10.1080/00031305.2017.1375986]. The Dockerfile included with the project can also be used to launch a RStudio server locally independent of binder which is especially useful when more computational resources are required.

  - [`repo2docker`](https://repo2docker.readthedocs.io/en/latest/config_files.html) (`install.R`, `DESCRPTION`, `runtime.txt`) based on the Rocker images [@nuest]
- GPU [@noamross, @cboettig]
- Gigantum stack

@@

- [RCloud](https://github.com/att/rcloud/tree/master/docker) ?

>>>>

## Common or public work environments

The fact that Docker images are portable and well defined make them useful when more than one person needs access to the same computing environment.
This is even more useful when some of the users do not have the expertise to create such an environment themselves, and when these environments can be run in public or shared infrastructure.

[`holepunch`](https://github.com/karthik/holepunch) is an R package that was designed to make sharing work environments accessible to novice R users based on _Binder_.
The [Binder project](https://mybinder.readthedocs.io/en/latest/), maintained by the team behind Jupyter, makes it possible for users to create and share their computing environments with others \citep{jupyter_binder_2018}.
A _BinderHub_ allows anyone with access to a web browser and an internet connection to launch a temporary instance of these custom environments and execute any workflows contained within.
From a reproducibility standpoint, Binder makes it exceedingly easy to compile a paper, visualize data, and run small examples from papers or tutorials without the need for any local installation.
To set up Binder for a project, a user typically starts at an instance of a BinderHub and passes the location of a repository with a workspace, e.g., a hosted Git repository, or a data repository like Zenodo.
Binder's core internal tool is [repo2docker](https://repo2docker.readthedocs.io/en/latest/config_files.html).
It deterministically builds a Docker image by parsing the contents of a repository, e.g. project dependency configurations or simple configuration files.
In the most powerful case, `repo2docker` builds a given Dockerfile.
While this approach works well for most run of the mill Python projects, it is not so seamless for R projects.
For any R projects that use the Tidyverse suite [citation in review, will add by the time this gets submitted], the time and resources required to build all dependencies from source can often time out before completion, making it frustrating for the average R user.
`holepunch` removes some of these limitations by leveraging Rocker images that contain the Tidyverse along special Jupyter dependencies, and only installs additional packages from CRAN and Bioconductor that are not already part of these images.
It short cicuits the configuration file parsing in `repo2docker` and starts with the Binder/Tidyverse base images, which eliminates a large part of the build time and in most cases results in a binder instance launching within a minute.
`holepunch` as a side effect also creates a `DESCRIPTION` file which then turns any project into a research compendium \citep{marwick_packaging_2018}.
The Dockerfile included with the project can also be used to launch a RStudio server locally independent of Binder which is especially useful when more or special computational resources are required.

@@

- GPU environments [@noamross, @cboettig]
- [RCloud](https://github.com/att/rcloud/tree/master/docker) ?
